---
title: "Investigating the Enron Scandal"
output: html_notebook
---
# 1. First, we need to install and call necessary libraries
```{r}
#install.packages(c('igraph','threejs'))

# For creating network graphs
library(igraph)  
library(visNetwork)
# For creating interactive, 3D visualisations of a network
```


# 2. Read in the email dataset from the data folder
```{r}
# Read the data
data <- read.csv('enron_emails.csv')

# First 5 rows
head(data)

# Notice that it's quite simply two columns: From, and  To
```


# 3. Create a graph from the data
```{r}
# Create a directed graph from the data
g <- graph_from_edgelist(as.matrix(data), directed = T)
```


# 4. Investigate our Enron email network

```{r}
# See the vertices/nodes
V(g)

# Number of nodes/vertices/Enron employees in the network
gorder(g)

# See the edges
E(g)

# Number of edges/emails
gsize(g)

# Check if our graph is directed
is.directed(g)

# Check if our network graph is weighted
is.weighted(g)
```


## Explore adjacent nodes/neighbours

```{r}
# We can see adjacent nodes using the 'neighbors' function

# Let's explore Gregg  Whalley's neighbours - first, all emails that were sent to him, i.e. mode = 'in'
neighbors(g, 'greg.whalley', mode = c('in'))

# How about emails he sent?
neighbors(g, 'greg.whalley', mode = c('out'))

# How about all emails?
neighbors(g, 'greg.whalley', mode = c('all'))

```


## Explore degrees of the nodes in the network

```{r}
# Find the degrees of each node --  in other words, number  of people each employee has sent emails to or received emails from
degree(g)

# Sort this  list in decreasing  order
sort(degree(g), decreasing = T)
```



## Explore cliques and communities in the network -- i.e. completed networks

```{r}
# Let's now find all cliques in the network

# If the 'min' argument isn't specified, it will include individuals as cliques of their own
cliques(g, min = 6)

# Find the largest  present clique in the Enron email network
largest_cliques(g)

```


## Plot the largest clique present in the network as an arcplot

```{r}
# Assign a variable to larget cliques
lclique <- largest_cliques(g)

# Create a new subgraph only containing the vertices of the largest clique we found above
subg <- induced_subgraph(g, lclique[[1]])

# Plot an 'arcplot' of this subgraph
plot(subg,
     vertex.label.color = "black",
     vertex.label.cex = 0.9,
     vertex.size = 0,
     edge.color = 'gray28',
     main = "Network of Largest Clique at Enron",
     layout = layout.circle(subg)
)
```


# 5. Find and visualise important nodes.

## Using 3 measures of centrality:
    1) Betweenness centrality
    2) Eigenvector centrality
    3) Pagerank


## Betweenness centrality
The more shortest paths that go through the node, the more important it is

```{r}
library(visNetwork)
```


```{r}
visIgraph(g)
```


```{r}
data <- toVisNetworkData(g)

# Create numerical vector of various centrality measures --  betweenness, eignecentrality, and pagerank.

data$nodes$value <- as.numeric(betweenness(g))
data$nodes$title <- data$nodes$label
data$nodes$label <- NULL

visNetwork(nodes = data$nodes, edges = data$edges) %>% 
  visIgraphLayout(randomSeed = 123) %>%
  visNodes(scaling = list(min = 30, max = 100), color = list(background = "lightblue", highlight = "yellow")) %>%
  visEdges(color = list(color = "cornflowerblue", highlight = "red"))
```




## Eigenvector centrality
The more well-connected the node is to other well-connected nodes, the more important it is.

```{r}
data <- toVisNetworkData(g)

# Create numerical vector of various centrality measures --  betweenness, eignecentrality, and pagerank.

data$nodes$value <- as.numeric(eigen_centrality(g)$vector)
data$nodes$title <- data$nodes$label
data$nodes$label <- NULL

visNetwork(nodes = data$nodes, edges = data$edges) %>% 
  visIgraphLayout(randomSeed = 123) %>%
  visNodes(scaling = list(min = 30, max = 100), color = list(background = "lightblue", highlight = "yellow")) %>%
  visEdges(color = list(color = "cornflowerblue", highlight = "red"))
```


## Centrality according to the PageRank algorithm
Node importance is dictated by:
  - Number of nodes it's connected to
  - More importantly, the 'quality' of those nodes.

```{r}
data <- toVisNetworkData(g)

data$nodes$value <- as.numeric(page_rank(g)$vector)
data$nodes$title <- data$nodes$label
data$nodes$label <- NULL

visNetwork(nodes = data$nodes, edges = data$edges) %>% 
  visIgraphLayout(randomSeed = 123) %>%
  visNodes(scaling = list(min = 30, max = 100), color = list(background = "lightblue", highlight = "yellow")) %>%
  visEdges(color = list(color = "cornflowerblue", highlight = "red"))
```




## Some useful insights:

1) Tracy Geaccone and Rod Hayslett are both the largest nodes in the network, according to Eigenvector centrality. They are also part of the largest clique, as we discovered above.  

According to the 'Report of Investigation of Enron Corporation and Related Entities Regarding Federal Tax and Compensation Issues, and Policy Recommendations, Volume III: Appendices C & D' (https://books.google.co.uk/books?id=UggNpmlzYFYC&pg=PA150&lpg=PA150&dq=tracy+geaccone+enron&source=bl&ots=X0HQc5OZci&sig=ACfU3U0Fwm1j8cHe_v8Hh9btALrsN7xGyQ&hl=en&sa=X&ved=2ahUKEwjFjJPutujgAhWpQhUIHSOYBUoQ6AEwA3oECAcQAQ#v=onepage&q=tracy%20geaccone&f=false)

 -  Tracy Geaccone: Director of Consolidations in the ETS department
 -  Rod (Roderick) Hayslett: Managing Director, ETS Finance and Accounting

 Both in the department where the fraud took place!


 2) Also on the Eigenvector centrality visualisation, can you find:
    - Kenneth Lay, Founder, Chairman, Ex-CEO of Enron
    - Jeff  (Jeffrey) Skilling, CEO of Enron at the  time

 3) The PageRank visualisation doesn't seem to show massive differences in node importance - why is that?


Did you notice anything else in your investigation?




Interesting notebook on RPubs on Enron dataset: https://rpubs.com/Koba/enron-network-code. One of the plots at the bottom shows how cliques started forming as time went by, and we got closer to the time of the scandal.